### 2021-04-05 My-Papers/SpecHunter/evaluation.tex
minors

\section{Evaluation}

\subsection{RQs}
\begin{itemize}
\item How effective is SH in finding src/sin/san?  % I am really really worried about this
\item How efficient is SH?
\item Interaction user-study % possible?
\end{itemize}

% why is this meaningful? (with/without comparison)
% at least mention comparison

\subsection{Quantitative}

\subsection{Qualitative}

\paragraph{Discussions}
How well are propagations working? (Quant, Qualit) .
How stable is SpecHunter over multiple iterations?
\paragraph{Limitations}

\paragraph{Threats to Validity}